Population structure and phenotypic variation of \emph{Sclerotinia sclerotiorum} from dry bean (\emph{Phaseolus vulgaris}) in the United States

### 2017-11-10 doc/manuscript/manuscript.tex
minor formatting for labels [skip ci]

A total of 366 isolates were collected from 2003 to 2012 (except 2006
and 2011) from diseased dry bean plants in 11 states in the United
States as well as Australia, France, and Mexico (Table
\ref{tab:isolate-table}). With the 11 loci used in the analyses (Table
\ref{tab:locus-stats}), we observed a total of 165 MLHs (215 with 16
loci). These 11 loci are found on 7 chromosomes in the \emph{S.
sclerotiorum} genome with a minimum distance of 55Kbp between two loci
on the same chromosome. Over 50\% of the isolates came from four states,
MI (62), ND (60), WA (59), NE (47). Four regions had fewer than 10
isolates, Australia (6), WI (2), NY (1), ID (1). We observed 87 MCGs,
the most abundant of which (`MCG 5') was represented by 73 isolates over
37 MLHs (Fig \ref{biggraph}A,C).

@@

Agressiveness of the isolates ranged from 1.4 to 7.9 with a mean of 5.02
and median of 4.85. The group mean averages were 4.88, 5.13, and 5.19
for Region, MCG, and MLH, respectively. A strip plot showing the
distribution of severity across these three variables simultaneously can
be seen in figure \ref{mcg-mlh-region-aggressiveness}. Our assessment of
aggressiveness in association with Region showed a significant effect
(\emph{P} \textless{} 1.00e\textsuperscript{-4}), with means that ranged
from 5.8 (MN) to 4.0 (CA) (Fig. \ref{aggressiveness-boxplot}, Table
\ref{tab:region-aggressiveness}). MCGs also showed a significant effect
(\emph{P} \textless{} 0.001), with means that ranged from 6.0 (`MCG 44')
to 4.6 (`MCG 49'; Table \ref{tab:MCG-aggressiveness}). We additionally
found a significant effect for MLHs (\emph{P} \textless{} 0.001), with
means that ranged from 6.0 (`MLH 78') to 4.3 (`MLH 140') (Table
\ref{tab:MLG-aggressiveness}).

@@

In our analysis, we found 165 MLHs with 70 singletons and 87 MCGs with
43 singletons (Fig. \ref{biggraph}A,B) where the eight most abundant
MCGs represented \textgreater{} 51\% of the data over 11 Regions, and
all years except for 2012. Our network-based approach to correlating
MLHs with MCGs revealed a large and complex network (Fig.
\ref{biggraph}, Table \ref{tab:mlg-table}). Community analysis showed 51
communities, 15 of which consisted of a single MLH unconnected with any
other community indicating that just 9.09\% of the 165 MLHs are unable
to cross with any other MLH in this data set (Fig. \ref{fullgraph}). The
three communities with the most members contained eight of the 10 most
abundant MCGs. Comparing these communities with Bruvo's genetic distance
showed an average distance of 0.451 among communities and an average
distance of 0.437 within communities, which were not significantly
different. When we assessed the number of times two different MLHs that
are in the same MCG, considering these as potential heterothallic
pairings that could result in sexual recombination, we found an average
of 14.3 potential heterothallic parings per MLH. Representing just four
isolates, `MLH 75' had 57 neighbors that shared the same mycelial
compatibility group (Fig \ref{biggraph}, \ref{fullgraph}). Overall,
there was no clear pattern to the association between MLH and MCGs.

@@

With the exception of the WA Region, populations that were sampled
across several years appeared to be relatively stable over time with
overlapping distributions in the DAPC (i.e.~NE and MI, Fig.
\ref{DAPC-RY}). DAPC is based on the principal components of allele
counts (Jombart et al., 2010). Unlike Bruvo's distance, this does not
take into account the magnitude of the difference between alleles, which
could inflate the distance measure in the presence of private alleles
(Bruvo et al., 2004). While we found no evidence of private alleles in
the Mexico and CA isolates, we did find that the alleles driving the
first axis in figure \ref{DAPC}A (alleles 174, 256, and 372 in loci 7-2,
8-3, and 9-2, respectively) were overrepresented in Mexico (where
\textgreater{}75\% of the alleles came from the region). However, all
three of these alleles, i) conform to the expected stepwise mutation
model (Bruvo et al., 2004) and ii) are at or near the extremes of the
total range (except for allele 372 at locus 9-2). Moreover, the fact
that we find three alleles at three independent loci segregating for
Mexico suggests that the pattern separating these populations from the
others was not an artifact. We believe that the differences in
populations observed from Mexico may be due to differences in climate
that allow greater diversification via sexual outcrossing.

@@

When we tested for sexual reproduction, we were unable to find evidence
for it in any region except for Australia and CA. While the Australia
population had a non-significant value of \(\bar{r}_d\)---which would
suggest that we cannot reject the null hypothesis of random mating---the
sample size was insufficient from which to draw conclusions (Milgroom,
1996; Agapow \& Burt, 2001). The low value of \(\bar{r}_d\) in the CA
population may represent sexual reproduction, but we can see in Figure
\ref{DAPC-RY} that there is differentiation by year. Thus, this could
also be an artifact of sampling two different populations, which is
known to reduce the value of \(\bar{r}_d\) (Prugnolle \& de Meeûs,
2010).

@@

The authors would like to thank Rebecca Higgens for technical support in
generating the data for the MCG assessment, aggressiveness ratings, and
genotyping; and for providing valuable insights into the historical
context of the data collection and curation. We would also like to thank
Denita Hadziabdic and two other anonymous reviewers for their valuable
comments and insights that improved the quality of the manuscript.

\subsection*{Conflicts of Interest}\label{conflicts-of-interest}
\addcontentsline{toc}{subsection}{Conflicts of Interest}

>>>>

A total of 366 isolates were collected from 2003 to 2012 (except 2006
and 2011) from diseased dry bean plants in 11 states in the United
States as well as Australia, France, and Mexico (Table
\ref{tab:isolate-table}). With the 11 loci used in the analyses (Table
\ref{tab:locus-stats}), we observed a total of 165 MLHs (215 with 16
loci). These 11 loci are found on 7 chromosomes in the \emph{S.
sclerotiorum} genome with a minimum distance of 55Kbp between two loci
on the same chromosome. Over 50\% of the isolates came from four states,
MI (62), ND (60), WA (59), NE (47). Four regions had fewer than 10
isolates, Australia (6), WI (2), NY (1), ID (1). We observed 87 MCGs,
the most abundant of which (`MCG 5') was represented by 73 isolates over
37 MLHs (Fig. \ref{biggraph}A,C).

@@

Agressiveness of the isolates ranged from 1.4 to 7.9 with a mean of 5.02
and median of 4.85. The group mean averages were 4.88, 5.13, and 5.19
for Region, MCG, and MLH, respectively. A strip plot showing the
distribution of severity across these three variables simultaneously can
be seen in Fig. \ref{mcg-mlh-region-aggressiveness}. Our assessment of
aggressiveness in association with Region showed a significant effect
(\emph{P} \textless{} 1.00e\textsuperscript{-4}), with means that ranged
from 5.8 (MN) to 4.0 (CA) (Fig. \ref{aggressiveness-boxplot}, Table
\ref{tab:region-aggressiveness}). MCGs also showed a significant effect
(\emph{P} \textless{} 0.001), with means that ranged from 6.0 (`MCG 44')
to 4.6 (`MCG 49'; Table \ref{tab:MCG-aggressiveness}). We additionally
found a significant effect for MLHs (\emph{P} \textless{} 0.001), with
means that ranged from 6.0 (`MLH 78') to 4.3 (`MLH 140') (Table
\ref{tab:MLG-aggressiveness}).

@@

In our analysis, we found 165 MLHs with 70 singletons and 87 MCGs with
43 singletons (Fig. \ref{biggraph}A,B) where the eight most abundant
MCGs represented \textgreater{} 51\% of the data over 11 Regions, and
all years except for 2012. Our network-based approach to correlating
MLHs with MCGs revealed a large and complex network (Fig.
\ref{biggraph}, Table \ref{tab:mlg-table}). Community analysis showed 51
communities, 15 of which consisted of a single MLH unconnected with any
other community indicating that just 9.09\% of the 165 MLHs are unable
to cross with any other MLH in this data set (Fig. \ref{fullgraph}). The
three communities with the most members contained eight of the 10 most
abundant MCGs. Comparing these communities with Bruvo's genetic distance
showed an average distance of 0.451 among communities and an average
distance of 0.437 within communities, which were not significantly
different. When we assessed the number of times two different MLHs that
are in the same MCG, considering these as potential heterothallic
pairings that could result in sexual recombination, we found an average
of 14.3 potential heterothallic parings per MLH. Representing just four
isolates, `MLH 75' had 57 neighbors that shared the same mycelial
compatibility group (Fig. \ref{biggraph}, \ref{fullgraph}). Overall,
there was no clear pattern to the association between MLH and MCGs.

@@

With the exception of the WA Region, populations that were sampled
across several years appeared to be relatively stable over time with
overlapping distributions in the DAPC (i.e.~NE and MI, Fig.
\ref{DAPC-RY}). DAPC is based on the principal components of allele
counts (Jombart et al., 2010). Unlike Bruvo's distance, this does not
take into account the magnitude of the difference between alleles, which
could inflate the distance measure in the presence of private alleles
(Bruvo et al., 2004). While we found no evidence of private alleles in
the Mexico and CA isolates, we did find that the alleles driving the
first axis in Fig. \ref{DAPC}A (alleles 174, 256, and 372 in loci 7-2,
8-3, and 9-2, respectively) were overrepresented in Mexico (where
\textgreater{}75\% of the alleles came from the region). However, all
three of these alleles, i) conform to the expected stepwise mutation
model (Bruvo et al., 2004) and ii) are at or near the extremes of the
total range (except for allele 372 at locus 9-2). Moreover, the fact
that we find three alleles at three independent loci segregating for
Mexico suggests that the pattern separating these populations from the
others was not an artifact. We believe that the differences in
populations observed from Mexico may be due to differences in climate
that allow greater diversification via sexual outcrossing.

@@

When we tested for sexual reproduction, we were unable to find evidence
for it in any region except for Australia and CA. While the Australia
population had a non-significant value of \(\bar{r}_d\)---which would
suggest that we cannot reject the null hypothesis of random mating---the
sample size was insufficient from which to draw conclusions (Milgroom,
1996; Agapow \& Burt, 2001). The low value of \(\bar{r}_d\) in the CA
population may represent sexual reproduction, but we can see in Fig.
\ref{DAPC-RY} that there is differentiation by year. Thus, this could
also be an artifact of sampling two different populations, which is
known to reduce the value of \(\bar{r}_d\) (Prugnolle \& de Meeûs,
2010).

@@

The authors would like to thank Rebecca Higgens for technical support in
generating the data for the MCG assessment, aggressiveness ratings, and
genotyping; and for providing valuable insights into the historical
context of the data collection and curation.

We would also like to thank Denita Hadziabdic and two other anonymous
reviewers for their valuable comments and insights that improved the
quality of the manuscript. \#\# Conflicts of Interest \{-\}
